---
title: "<center> <h1>DATA VALIDATION</h1> </center>"
author: "<center> <h3> Cecilia Siliato </h3> </center>"
output:
  html_document: default
  word_document: default
  pdf_document: default
subtitle: <center> <h2> Monitoring Arm Activity in People with Neuromuscular Impairments
  </h2> </center>
number_sections: yes
---

##DATA DESCRIPTION

Here, there are two databases that contains data from an Arm Tracker System which is a wearable system based on inertial measurement units (IMUs) that can measure arm movement during daily life. The two databases refers to: Healthy data and data that comes from arm movement in a person with Becker muscular dystrophy.

The data correspond to approximately 6 hours of arms movement and the frequency is of 20 milliseconds. There are 5 sensors located on torso, the 2 elbows and 2 wrists. By the moment, we will focus our efforts on understanding the following variable types:

1. Timestamp: Integer number indicating in which millisecond we are (starting from 20).

2. Orientation (quaternion): We have 2 types of orientations; the first is the relative orientation of each sensor and the world and the second one is the relative orientation of each arm sensor and the torso one. The data of each one of these variables is a quaternion.

3. Position (x,y,z)

4. Angles

5. Acceleration



##DESCRIPTION OF THE PROCEDURE

First of all, we must comment that an R package (which is in continuous development) has been created where useful functions for the handling of data from Arm Tracker are included and modified periodically.

```{r, warning=FALSE, message=FALSE}
#Installing package
install.packages("c:/TFM/Code/armtrack_0.0.0.9000.tar.gz", repos = NULL, type = "source")
library(armtrack)
```


### 1. Reading Data

In order to make the data reading more agile, we have developed a function called <i>at.read</i>. This function reads data coming from Arm Tracker in '.mat' format and it performs some little changes on the database in order to work better with quaternions and position vectors.

The only parameter that must be passed is the path from where we will take the database:

```{r, warning=FALSE, message=FALSE}
Hdata<-"c:/TFM/Data/HealthyData.mat"
Hdata<-at.read(Hdata)
```

```{r, warning=FALSE, message=FALSE}
Bdata<-"c:/TFM/Data/BMDData.mat"
Bdata<-at.read(Bdata)
```

Now we have a dataset with 54 variables including all the types commented before. 

### 2. Missing data: IMPORTANT?

With the purpose of performing the missing data analysis, we have developed a simple R function (<i>at.missing</i>) that follows the procedure:

1. Inserts the missing rows on the input dataset (when there is a difference on timestamp bigger than 20 between two consecutives rows, there is a missing rows).

2. Check if the % of missing cells is higher than a given parameter (by default, 5).

    2.1. If it is higher than the value given, it prints the message: "The x% of cells are missings. You should perform a deeper study." 

    2.2. If it is lower than the value given (most of the cases), it imputs for each missing cell the next non-missing value of that variable. The function allways informs the % of cells that were missing initially.

In our case:

```{r, warning=FALSE, message=FALSE}
#Imputing missing values
Hd_nomis<-at.missing(Hdata)
Bd_nomis<-at.missing(Bdata)
```

In order to perform the other part of the analysis, we will create a dataset with both data (healthy and unhealthy) with a factor variable indicating if the data corresponds to the healthy person (Healthy) or the person with Becker muscular distrophy (BMDistrophy).

```{r, warning=FALSE, message=FALSE}
Hd_nomis$Type<-"Healthy"
Bd_nomis$Type<-"BMDistrophy"
dd<-rbind.data.frame(Hd_nomis,Bd_nomis)
```

### 3. Univariant analysis: Validation of each variable

*1. QUATERNIONS*

<!-- regular html http://www.tobynorris.com/work/prog/csharp/quatview/help/orientations_and_quaternions.htm --> 

The quaternions are 4-component vectors. The first one is the real one and the other 3 are imaginary components (i, j and k). The quaternions show us the orientation of the sensor in relation to the world or to other sensor. We can obtain the rotation angle from the first component; but by the moment we are not going to centre our efforts in it.

```{r,warning=F, message=F,echo=F}
HD<-subset(Hd_nomis, select=c("Timestamp","ang","LElbAngle","RElbAngle","LSAngle","RSAngle","RHandPos_2","RHandPos_3","RHandPos_4","LHandPos_2","LHandPos_3","LHandPos_4","RElbPos_2","RElbPos_3","RElbPos_4","LElbPos_2","LElbPos_3","LElbPos_4","Type"))
BMD<-subset(Bd_nomis, select=c("Timestamp","ang","LElbAngle","RElbAngle","LSAngle","RSAngle","RHandPos_2","RHandPos_3","RHandPos_4","LHandPos_2","LHandPos_3","LHandPos_4","RElbPos_2","RElbPos_3","RElbPos_4","LElbPos_2","LElbPos_3","LElbPos_4","Type"))
TD<-subset(dd, select=c("Timestamp","ang","LElbAngle","RElbAngle","LSAngle","RSAngle","RHandPos_2","RHandPos_3","RHandPos_4","LHandPos_2","LHandPos_3","LHandPos_4","RElbPos_2","RElbPos_3","RElbPos_4","LElbPos_2","LElbPos_3","LElbPos_4","Type"))
rm(Hdata,Bdata,Bd_nomis,Hd_nomis,dd)
```
*2. Angles*

In order to validate the other angles that are present on the dataset, we can see the histogram of them:
    
```{r, warning=FALSE, message=FALSE,echo=FALSE}
if (!require("ggplot2"))
  {
  install.packages("ggplot2")
  library(ggplot2)
}
if (!require("gridExtra"))
{
  install.packages("gridExtra")
  library(gridExtra)
}
if (!require("grid"))
{
  install.packages("grid")
  library(grid)
}
if (!require("plotly"))
{
  install.packages("plotly")
  library(plotly)
}
get_legend<-function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
}

p1<-ggplot(TD, aes(LElbAngle, color=Type)) +  geom_histogram(alpha = 0.5, aes(y = ..density..), position = 'identity', fill = "white", binwidth=1) + scale_x_continuous(name ="Left",breaks=seq(0,150,10),limits=c(0,150),labels=c("0","","","30","","","60","","","90","","","120","","","150"))+theme(legend.position = "none")+ scale_y_continuous(name="",breaks=seq(0,0.0425,0.0025),limits=c(0,0.0425))

p2<-ggplot(TD, aes(RElbAngle, color=Type)) + geom_histogram(alpha = 0.5, aes(y = ..density..), position = 'identity', fill = "white", binwidth=1) +scale_x_continuous(name ="Right",breaks=seq(0,150,10),limits=c(0,150),labels=c("0","","","30","","","60","","","90","","","120","","","150"))+theme(legend.position = "none")+ scale_y_continuous(name="",breaks=seq(0,0.0425,0.0025),limits=c(0,0.0425),  labels=rep("",18))

#legend<-get_legend(p1)
p1<-ggplotly(p1)
p2<-ggplotly(p2)
subplot(p1, p2)%>%
layout(title = list(text = "Elbow angle", y = 0.995))
```
    
```{r, warning=FALSE, message=FALSE,echo=FALSE}
p1<-ggplot(TD, aes(LSAngle, color=Type)) + 
  geom_histogram(alpha = 0.5, aes(y = ..density..), position = 'identity', fill = "white", binwidth=1) + scale_x_continuous(name ="Left",breaks=seq(0,150,10),limits=c(0,150),labels=c("0","","","30","","","60","","","90","","","120","","","150"))+ scale_y_continuous(name="",breaks=seq(0,0.036,0.002),limits=c(0,0.035))+theme(legend.position = "top")
p11<-p1+theme(legend.position = "none")
p2<-ggplot(TD, aes(RSAngle, color=Type)) + 
  geom_histogram(alpha = 0.5, aes(y = ..density..), position = 'identity', fill = "white", binwidth=1) + scale_x_continuous(name ="Right",breaks=seq(0,150,10),limits=c(0,150),labels=c("0","","","30","","","60","","","90","","","120","","","150"))+ scale_y_continuous(name="", breaks=seq(0,0.036,0.002),limits=c(0,0.035), labels=rep("",19))+ theme(legend.position="none")

p11<-ggplotly(p11)
p2<-ggplotly(p2)
subplot(p11, p2)%>%
layout(title = list(text = "Upper arm and body angle", y = 0.995))
```
    
Note that these angles are directly expressed in degrees, not radians. Neither the Elbow angle nor the upper arm and body angle arrive never to 180ยบ. This is because the guy who was using the T-Shirt that collect the data was only working on the office and it would be strange to arrive to these degrees.

If we compare Healthy and Becker Distrophy Data, we can see more frequency and less dispersion in Becker Distrophy data. This makes sense regarding to the fact that people with Becker Distrophy has less mobility.

Finally we have compared the torso inclination angle between Healthy and Becker Distrophy data:

```{r}
p1<-ggplot(TD, aes(ang, color=Type)) + 
  geom_histogram(alpha = 0.5, aes(y = ..density..), position = 'identity', fill = "white", binwidth=1) + scale_x_continuous(name ="Torso angle",breaks=seq(0,180,10),limits=c(0,180),labels=c("0","","","30","","","60","","","90","","","120","","","150","","","180"))+scale_y_continuous(name="",breaks=seq(0,0.036,0.002),limits=c(0,0.035))+theme(legend.position = "none")
p1<-ggplotly(p1)
p1
```

People with Becker muscular Distrophy  has less movement on their extremities but more movement on the torso and this histogram shows us exactly this detail. We can see more frequency and less variance in healthy data than in BMD data. This is one of the few cases where we are going to see this situation.



*3. Position variables*

There are 4 position data reffering to (right and left) elbow position and (right and left) hand position. On the '.mat' database, these variables are represented as quaternions but on the preprocess performed, we realised that there is allways a component of these quaternions that was with value equal (or very similar) to 0. So, we have converted each position quaternion in 3 numeric variables corresponding to the 3 axes x, y and z. Therefore, we have finally $3 \cdot 4=12$ variables reffering to 4 different positions.

We are going to see density plots of the planes XY, YZ and XZ in order to see where are concentrated the biggest quantity of movements. But first, we reduce the data frequency (from 50 records per second to 5 records per second) to optimize the computation time:

```{r, warning=FALSE, message=FALSE,echo=FALSE}
se<-seq(0,nrow(HD),10)
HD1<-HD[rownames(HD) %in% se,]
se<-seq(0,nrow(BMD),10)
BMD1<-BMD[rownames(BMD) %in% se,]
se<-seq(0,nrow(TD),10)
TD1<-TD[rownames(TD) %in% se,]
rm(se)
```


```{r, warning=FALSE, message=FALSE,echo=FALSE}
#Plano XY:
par(mfrow=c(2,3))
## Use densCols() output to get density at each point
x <- densCols(BMD1$LElbPos_2,BMD1$LElbPos_3, colramp=colorRampPalette(c("black", "white")))
BMD1$dens <- col2rgb(x)[1,] + 1L
## Map densities to colors
cols <-  colorRampPalette(c("#000099", "#00FEFF", "#45FE4F", "#FCFF00", "#FF9400", "#FF3100"))(256)
BMD1$col <- cols[BMD1$dens]
## Plot it, reordering rows so that densest points are plotted on top
plot(LElbPos_3~LElbPos_2, data=BMD1[order(BMD1$dens),], pch=20, col=col, cex=2, xlab="x", ylab="y", main="BMDistrophy: XY",xlim=c(-300,150),ylim=c(0,450))
#Plano YZ:
## Use densCols() output to get density at each point
x <- densCols(BMD1$LElbPos_3,BMD1$LElbPos_4, colramp=colorRampPalette(c("black", "white")))
BMD1$dens <- col2rgb(x)[1,] + 1L
## Map densities to colors
cols <-  colorRampPalette(c("#000099", "#00FEFF", "#45FE4F", "#FCFF00", "#FF9400", "#FF3100"))(256)
BMD1$col <- cols[BMD1$dens]
## Plot it, reordering rows so that densest points are plotted on top
plot(LElbPos_4~LElbPos_3, data=BMD1[order(BMD1$dens),], pch=20, col=col, cex=2, xlab="y", ylab="z", main="BMDistrophy: YZ",xlim=c(0,450),ylim=c(1000,1650))
#Plano XZ:
## Use densCols() output to get density at each point
x <- densCols(BMD1$LElbPos_2,BMD1$LElbPos_4, colramp=colorRampPalette(c("black", "white")))
BMD1$dens <- col2rgb(x)[1,] + 1L
## Map densities to colors
cols <-  colorRampPalette(c("#000099", "#00FEFF", "#45FE4F", "#FCFF00", "#FF9400", "#FF3100"))(256)
BMD1$col <- cols[BMD1$dens]
## Plot it, reordering rows so that densest points are plotted on top
plot(LElbPos_4~LElbPos_2, data=BMD1[order(BMD1$dens),], pch=20, col=col, cex=2, xlab="x", ylab="z", main="BMDistrophy: XZ",xlim=c(-300,150),ylim=c(1000,1650))
BMD1 <- subset(BMD1,select=-c(col,dens))

#Plano XY
## Use densCols() output to get density at each point
x <- densCols(HD1$LElbPos_2,HD1$LElbPos_3, colramp=colorRampPalette(c("black", "white")))
HD1$dens <- col2rgb(x)[1,] + 1L
## Map densities to colors
cols <-  colorRampPalette(c("#000099", "#00FEFF", "#45FE4F","#FCFF00", "#FF9400", "#FF3100"))(256)
HD1$col <- cols[HD1$dens]
## Plot it, reordering rows so that densest points are plotted on top
plot(LElbPos_3~LElbPos_2, data=HD1[order(HD1$dens),], pch=20, col=col, cex=2, xlab="x", ylab="y", main="Healthy: XY",xlim=c(-300,150),ylim=c(0,450))
#Plano YZ:
## Use densCols() output to get density at each point
x <- densCols(HD1$LElbPos_3,HD1$LElbPos_4, colramp=colorRampPalette(c("black", "white")))
HD1$dens <- col2rgb(x)[1,] + 1L
## Map densities to colors
cols <-  colorRampPalette(c("#000099", "#00FEFF", "#45FE4F","#FCFF00", "#FF9400", "#FF3100"))(256)
HD1$col <- cols[HD1$dens]
## Plot it, reordering rows so that densest points are plotted on top
plot(LElbPos_4~LElbPos_3, data=HD1[order(HD1$dens),], pch=20, col=col, cex=2, xlab="y", ylab="z", main="Healthy: YZ",xlim=c(0,450),ylim=c(1000,1650))
#Plano XZ:
## Use densCols() output to get density at each point
x <- densCols(HD1$LElbPos_2,HD1$LElbPos_4, colramp=colorRampPalette(c("black", "white")))
HD1$dens <- col2rgb(x)[1,] + 1L
## Map densities to colors
cols <-  colorRampPalette(c("#000099", "#00FEFF", "#45FE4F","#FCFF00", "#FF9400", "#FF3100"))(256)
HD1$col <- cols[HD1$dens]
## Plot it, reordering rows so that densest points are plotted on top
plot(LElbPos_4~LElbPos_2, data=HD1[order(HD1$dens),], pch=20, col=col, cex=2, xlab="x", ylab="z", main="Healthy: XZ",xlim=c(-300,150),ylim=c(1000,1650))
HD1 <- subset(HD1,select=-c(dens,col))
title("Left Elbow position", line=-1.3,outer=T,cex.main=1.5)

#Plano XY:
par(mfrow=c(2,3))
## Use densCols() output to get density at each point
x <- densCols(BMD1$RElbPos_2,BMD1$RElbPos_3, colramp=colorRampPalette(c("black", "white")))
BMD1$dens <- col2rgb(x)[1,] + 1L
## Map densities to colors
cols <-  colorRampPalette(c("#000099", "#00FEFF", "#45FE4F", "#FCFF00", "#FF9400", "#FF3100"))(256)
BMD1$col <- cols[BMD1$dens]
## Plot it, reordering rows so that densest points are plotted on top
plot(RElbPos_3~RElbPos_2, data=BMD1[order(BMD1$dens),], pch=20, col=col, cex=2, xlab="x", ylab="y", main="BMDistrophy: XY",xlim=c(-400,200),ylim=c(-450,50))
#Plano YZ:
## Use densCols() output to get density at each point
x <- densCols(BMD1$RElbPos_3,BMD1$RElbPos_4, colramp=colorRampPalette(c("black", "white")))
BMD1$dens <- col2rgb(x)[1,] + 1L
## Map densities to colors
cols <-  colorRampPalette(c("#000099", "#00FEFF", "#45FE4F", "#FCFF00", "#FF9400", "#FF3100"))(256)
BMD1$col <- cols[BMD1$dens]
## Plot it, reordering rows so that densest points are plotted on top
plot(RElbPos_4~RElbPos_3, data=BMD1[order(BMD1$dens),], pch=20, col=col, cex=2, xlab="y", ylab="z", main="BMDistrophy: YZ",xlim=c(-400,70),ylim=c(1000,1650))
#Plano XZ:
## Use densCols() output to get density at each point
x <- densCols(BMD1$RElbPos_2,BMD1$RElbPos_4, colramp=colorRampPalette(c("black", "white")))
BMD1$dens <- col2rgb(x)[1,] + 1L
## Map densities to colors
cols <-  colorRampPalette(c("#000099", "#00FEFF", "#45FE4F", "#FCFF00", "#FF9400", "#FF3100"))(256)
BMD1$col <- cols[BMD1$dens]
## Plot it, reordering rows so that densest points are plotted on top
plot(RElbPos_4~RElbPos_2, data=BMD1[order(BMD1$dens),], pch=20, col=col, cex=2, xlab="x", ylab="z", main="BMDistrophy: XZ",xlim=c(-400,200),ylim=c(1000,1650))
BMD1 <- subset(BMD1,select=-c(col,dens))


## Use densCols() output to get density at each point
x <- densCols(HD1$RElbPos_2,HD1$RElbPos_3, colramp=colorRampPalette(c("black", "white")))
HD1$dens <- col2rgb(x)[1,] + 1L
## Map densities to colors
cols <-  colorRampPalette(c("#000099", "#00FEFF", "#45FE4F","#FCFF00", "#FF9400", "#FF3100"))(256)
HD1$col <- cols[HD1$dens]
## Plot it, reordering rows so that densest points are plotted on top
plot(RElbPos_3~RElbPos_2, data=HD1[order(HD1$dens),], pch=20, col=col, cex=2, xlab="x", ylab="y", main="Healthy: XY",xlim=c(-400,200),ylim=c(-450,50))
#Plano YZ:
## Use densCols() output to get density at each point
x <- densCols(HD1$RElbPos_3,HD1$RElbPos_4, colramp=colorRampPalette(c("black", "white")))
HD1$dens <- col2rgb(x)[1,] + 1L
## Map densities to colors
cols <-  colorRampPalette(c("#000099", "#00FEFF", "#45FE4F","#FCFF00", "#FF9400", "#FF3100"))(256)
HD1$col <- cols[HD1$dens]
## Plot it, reordering rows so that densest points are plotted on top
plot(RElbPos_4~RElbPos_3, data=HD1[order(HD1$dens),], pch=20, col=col, cex=2, xlab="y", ylab="z", main="Healthy: YZ",xlim=c(-400,70),ylim=c(1000,1650))
#Plano XZ:
## Use densCols() output to get density at each point
x <- densCols(HD1$RElbPos_2,HD1$RElbPos_4, colramp=colorRampPalette(c("black", "white")))
HD1$dens <- col2rgb(x)[1,] + 1L
## Map densities to colors
cols <-  colorRampPalette(c("#000099", "#00FEFF", "#45FE4F","#FCFF00", "#FF9400", "#FF3100"))(256)
HD1$col <- cols[HD1$dens]
## Plot it, reordering rows so that densest points are plotted on top
plot(RElbPos_4~RElbPos_2, data=HD1[order(HD1$dens),], pch=20, col=col, cex=2, xlab="x", ylab="z", main="Healthy: XZ",xlim=c(-400,200),ylim=c(1000,1650))
HD1 <- subset(HD1,select=-c(dens,col))

title("Right Elbow position", line=-1.3,outer=T,cex.main=1.5)
```

As we expected, we can see more mobility on the healthy data elbows. Becker data has more red points (the positions are more frequents) and the movements are concentrated in a smaller space.

On the other hand, mainly in the plots in which the z axis appears, we can observe the rotation movement.

```{r, echo=F, warning=F, message=F}
######LEFT WRIST######
#Plano XY:
par(mfrow=c(2,3))
## Use densCols() output to get density at each point
x <- densCols(BMD1$LHandPos_2,BMD1$LHandPos_3, colramp=colorRampPalette(c("black", "white")))
BMD1$dens <- col2rgb(x)[1,] + 1L
## Map densities to colors
cols <-  colorRampPalette(c("#000099", "#00FEFF", "#45FE4F", "#FCFF00", "#FF9400", "#FF3100"))(256)
BMD1$col <- cols[BMD1$dens]
## Plot it, reordering rows so that densest points are plotted on top
plot(LHandPos_3~LHandPos_2, data=BMD1[order(BMD1$dens),], pch=20, col=col, cex=2, xlab="x", ylab="y", main="BMDistrophy: XY",xlim=c(-400,450),ylim=c(-250,650))
#Plano YZ:
## Use densCols() output to get density at each point
x <- densCols(BMD1$LHandPos_3,BMD1$LHandPos_4, colramp=colorRampPalette(c("black", "white")))
BMD1$dens <- col2rgb(x)[1,] + 1L
## Map densities to colors
cols <-  colorRampPalette(c("#000099", "#00FEFF", "#45FE4F", "#FCFF00", "#FF9400", "#FF3100"))(256)
BMD1$col <- cols[BMD1$dens]
## Plot it, reordering rows so that densest points are plotted on top
plot(LHandPos_4~LHandPos_3, data=BMD1[order(BMD1$dens),], pch=20, col=col, cex=2, xlab="y", ylab="z", main="BMDistrophy: YZ",xlim=c(-250,700),ylim=c(800,1750))
#Plano XZ:
## Use densCols() output to get density at each point
x <- densCols(BMD1$LHandPos_2,BMD1$LHandPos_4, colramp=colorRampPalette(c("black", "white")))
BMD1$dens <- col2rgb(x)[1,] + 1L
## Map densities to colors
cols <-  colorRampPalette(c("#000099", "#00FEFF", "#45FE4F", "#FCFF00", "#FF9400", "#FF3100"))(256)
BMD1$col <- cols[BMD1$dens]
## Plot it, reordering rows so that densest points are plotted on top
plot(LHandPos_4~LHandPos_2, data=BMD1[order(BMD1$dens),], pch=20, col=col, cex=2, xlab="x", ylab="z", main="BMDistrophy: XZ",xlim=c(-400,450),ylim=c(800,1750))
BMD1 <- subset(BMD1,select=-c(col,dens))

#Plano XY
## Use densCols() output to get density at each point
x <- densCols(HD1$LHandPos_2,HD1$LHandPos_3, colramp=colorRampPalette(c("black", "white")))
HD1$dens <- col2rgb(x)[1,] + 1L
## Map densities to colors
cols <-  colorRampPalette(c("#000099", "#00FEFF", "#45FE4F","#FCFF00", "#FF9400", "#FF3100"))(256)
HD1$col <- cols[HD1$dens]
## Plot it, reordering rows so that densest points are plotted on top
plot(LHandPos_3~LHandPos_2, data=HD1[order(HD1$dens),], pch=20, col=col, cex=2, xlab="x", ylab="y", main="Healthy: XY",xlim=c(-400,450),ylim=c(-250,650))
#Plano YZ:
## Use densCols() output to get density at each point
x <- densCols(HD1$LHandPos_3,HD1$LHandPos_4, colramp=colorRampPalette(c("black", "white")))
HD1$dens <- col2rgb(x)[1,] + 1L
## Map densities to colors
cols <-  colorRampPalette(c("#000099", "#00FEFF", "#45FE4F","#FCFF00", "#FF9400", "#FF3100"))(256)
HD1$col <- cols[HD1$dens]
## Plot it, reordering rows so that densest points are plotted on top
plot(LHandPos_4~LHandPos_3, data=HD1[order(HD1$dens),], pch=20, col=col, cex=2, xlab="y", ylab="z", main="Healthy: YZ",xlim=c(-250,700),ylim=c(800,1750))
#Plano XZ:
## Use densCols() output to get density at each point
x <- densCols(HD1$LHandPos_2,HD1$LHandPos_4, colramp=colorRampPalette(c("black", "white")))
HD1$dens <- col2rgb(x)[1,] + 1L
## Map densities to colors
cols <-  colorRampPalette(c("#000099", "#00FEFF", "#45FE4F","#FCFF00", "#FF9400", "#FF3100"))(256)
HD1$col <- cols[HD1$dens]
## Plot it, reordering rows so that densest points are plotted on top
plot(LHandPos_4~LHandPos_2, data=HD1[order(HD1$dens),], pch=20, col=col, cex=2, xlab="x", ylab="z", main="Healthy: XZ",xlim=c(-400,450),ylim=c(800,1750))
HD1 <- subset(HD1,select=-c(dens,col))

title("Left Wrist position", line=-1.3,outer=T,cex.main=1.5)


######RIGHT WRIST######

#Plano XY:
par(mfrow=c(2,3))
## Use densCols() output to get density at each point
x <- densCols(BMD1$RHandPos_2,BMD1$RHandPos_3, colramp=colorRampPalette(c("black", "white")))
BMD1$dens <- col2rgb(x)[1,] + 1L
## Map densities to colors
cols <-  colorRampPalette(c("#000099", "#00FEFF", "#45FE4F", "#FCFF00", "#FF9400", "#FF3100"))(256)
BMD1$col <- cols[BMD1$dens]
## Plot it, reordering rows so that densest points are plotted on top
plot(RHandPos_3~RHandPos_2, data=BMD1[order(BMD1$dens),], pch=20, col=col, cex=2, xlab="x", ylab="y", main="BMDistrophy: XY",xlim=c(-650,410),ylim=c(-700,250))
#Plano YZ:
## Use densCols() output to get density at each point
x <- densCols(BMD1$RHandPos_3,BMD1$RHandPos_4, colramp=colorRampPalette(c("black", "white")))
BMD1$dens <- col2rgb(x)[1,] + 1L
## Map densities to colors
cols <-  colorRampPalette(c("#000099", "#00FEFF", "#45FE4F", "#FCFF00", "#FF9400", "#FF3100"))(256)
BMD1$col <- cols[BMD1$dens]
## Plot it, reordering rows so that densest points are plotted on top
plot(RHandPos_4~RHandPos_3, data=BMD1[order(BMD1$dens),], pch=20, col=col, cex=2, xlab="y", ylab="z", main="BMDistrophy: YZ",xlim=c(-700,250),ylim=c(800,1850))
#Plano XZ:
## Use densCols() output to get density at each point
x <- densCols(BMD1$RHandPos_2,BMD1$RHandPos_4, colramp=colorRampPalette(c("black", "white")))
BMD1$dens <- col2rgb(x)[1,] + 1L
## Map densities to colors
cols <-  colorRampPalette(c("#000099", "#00FEFF", "#45FE4F", "#FCFF00", "#FF9400", "#FF3100"))(256)
BMD1$col <- cols[BMD1$dens]
## Plot it, reordering rows so that densest points are plotted on top
plot(RHandPos_4~RHandPos_2, data=BMD1[order(BMD1$dens),], pch=20, col=col, cex=2, xlab="x", ylab="z", main="BMDistrophy: XZ",xlim=c(-650,410),ylim=c(800,1850))
BMD1 <- subset(BMD1,select=-c(col,dens))


#Plano XY
## Use densCols() output to get density at each point
x <- densCols(HD1$RHandPos_2,HD1$RHandPos_3, colramp=colorRampPalette(c("black", "white")))
HD1$dens <- col2rgb(x)[1,] + 1L
## Map densities to colors
cols <-  colorRampPalette(c("#000099", "#00FEFF", "#45FE4F","#FCFF00", "#FF9400", "#FF3100"))(256)
HD1$col <- cols[HD1$dens]
## Plot it, reordering rows so that densest points are plotted on top
plot(RHandPos_3~RHandPos_2, data=HD1[order(HD1$dens),], pch=20, col=col, cex=2, xlab="x", ylab="y", main="Healthy: XY",xlim=c(-650,410),ylim=c(-700,250))
#Plano YZ:
## Use densCols() output to get density at each point
x <- densCols(HD1$RHandPos_3,HD1$RHandPos_4, colramp=colorRampPalette(c("black", "white")))
HD1$dens <- col2rgb(x)[1,] + 1L
## Map densities to colors
cols <-  colorRampPalette(c("#000099", "#00FEFF", "#45FE4F","#FCFF00", "#FF9400", "#FF3100"))(256)
HD1$col <- cols[HD1$dens]
## Plot it, reordering rows so that densest points are plotted on top
plot(RHandPos_4~RHandPos_3, data=HD1[order(HD1$dens),], pch=20, col=col, cex=2, xlab="y", ylab="z", main="Healthy: YZ",xlim=c(-700,250),ylim=c(800,1850))
#Plano XZ:
## Use densCols() output to get density at each point
x <- densCols(HD1$RHandPos_2,HD1$RHandPos_4, colramp=colorRampPalette(c("black", "white")))
HD1$dens <- col2rgb(x)[1,] + 1L
## Map densities to colors
cols <-  colorRampPalette(c("#000099", "#00FEFF", "#45FE4F","#FCFF00", "#FF9400", "#FF3100"))(256)
HD1$col <- cols[HD1$dens]
## Plot it, reordering rows so that densest points are plotted on top
plot(RHandPos_4~RHandPos_2, data=HD1[order(HD1$dens),], pch=20, col=col, cex=2, xlab="x", ylab="z", main="Healthy: XZ",xlim=c(-650,410),ylim=c(800,1850))
HD1 <- subset(HD1,select=-c(dens,col))

title("Right Wrist position", line=-1.3,outer=T,cex.main=1.5)
```

The comments here are very similar to the ones before. More variability in healthy data and there is (almost allways) more than one red point because of the bigger mobility of healthy people.
